Week 2 Peer Assignment for Reproducible Research
================================================
  
## Introduction

This assignment explores data collected from a personal activity monitoring
device. The device collected data at five minute intervals throughout the day
for a two month period and includes the number of steps taken in those five
minute  intervals.
  
  
### Load and preprocess the data

```{r}
activity <- read.csv('activity.csv')
activity$date <- as.Date(activity$date)
activity$steps <- as.numeric(as.character(activity$steps))
activity$interval <- as.numeric(as.character(activity$interval))

obs <- nrow(activity)
completeObs <- nrow(na.omit(activity))
missingObs <- obs - completeObs
```

There are `r obs` observations. `r missingObs` rows are missing observations.
  
### Histogram of the total number of steps taken each day

```{r}
total_steps_by_date <- aggregate(activity$steps, by=list(activity$date), FUN=sum)
hist(total_steps_by_date$x, main="Histogram of Total Steps Per Day", xlab="Total steps by day")
```    
  
### Calculate the mean and median total number of steps taken per day

For this calculation, we'll use the summary function which also includes
other information such a min, max, quartiles, and the number of NAs.

```{r}
summary(total_steps_by_date$x)
```

### Show the average number of steps per five minute interval for all days

```{r}
by_interval <- split(activity, activity$interval)
avg_by_interval <- sapply(by_interval, function(x) mean(x$steps, na.rm=TRUE))
plot(avg_by_interval, type="l",
     xlab="Interval",
     ylab="Average Number of Steps",
     main="Average Number of Steps Per Five Minute Interval",
     xaxt = "n")
xtic <- seq(0, length(avg_by_interval), 24)
xlabel <- seq(0, 2400, 200)
axis(1, at=xtic, labels=FALSE)
text(x=xtic, y=par("usr")[3]-15, labels=xlabel, srt=90, adj=1, xpd=TRUE)
```

The five minute interval which contains the maximum average number of steps
occurs at interval `r names(which.max(avg_by_interval))`.

### Filling in missing values

Of the `r obs` observations, `r missingObs` rows are missing values.



### Comparison of activity patterns between weekdays and weekends